---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ntsworkflow

The goal of ntsworkflow is to process non-target LC/GC-HRMS data and carry out simple data analysis. It is designed as a basic peak-picking and annotation algorithm which feeds into more specialized workflows for different tasks.

If you use the package or any derivative thereof, please cite the work.
To cite this work please use the following citation or run `citation("ntsworkflow")`.

> Kevin S. Jewell, Christian Dietrich, Toni Köppe, Franziska Thron, Arne Wick and Thomas A. Ternes (2023). 
> ntsworkflow: A Non-Target Screening Data Evaluation Tool. R package version 0.2.5.

### Peer-reviewed articles reporting the development of ntsworkflow

* K. S. Jewell, U. Kunkel, B. Ehlig, F. Thron, M. Schlüsener, C. Dietrich, A. Wick, and T. A. Ternes, *Rapid Communications in Mass Spectrometry*, **2019**, 34, e8541.
* T. Köppe, K. S. Jewell, C. Dietrich, A. Wick, and T. A. Ternes, *Water Research*, **2020**, 178, 115703.
* C. Dietrich, A. Wick, and T. A. Ternes, *Rapid Communications in Mass Spectrometry*, **2021**, e9206.

## Installation

R 4.5.0 and Java JRE and JDK (v1.8) (for rcdk package) are needed. Compilation of C++ code requires GCC v13.

The following guide uses [renv](https://rstudio.github.io/renv/) to manage the package dependencies. To start a new project use `renv::init(bioconductor = TRUE)`.

#### Install the Bioconductor package xcms

```{r, eval=FALSE}
renv::install("bioc::xcms")
```

#### Install Genform (optional)
The Windows .exe can be downloaded directly from Sourceforge. It must be put on the system path.

For Linux, the program needs to be compiled from source. Download the source files and in the folder with the .cpp files run the following:
`g++ main.cpp ms*.cpp -o genform`. This will compile the program and create the `genform` executable, which must be moved to the `~/bin` folder.

#### Install ntsworkflow

```{r, eval=FALSE}
renv::install("bafg-bund/ntsworkflow")
```

#### Installation on Ubuntu
Using `apt`, the packages `r-base`, `default-jre` and `default-jdk` are needed. Installing Rstudio server for WSL2 was done through the website instructions (deb package).

xcms requires mzR which in turn requires ncdf4 which needs the nc-config script to be installed. Prior to installing xcms you therefore need to install ncdf4. If, for some reason, using `renv::install`, `BiocManager::install` or `install.packages` does not work, you can use `apt` (`sudo apt install r-cran-ncdf4`) on Ubuntu. This installs ncdf4 into `/usr/lib/R/site-library`.

Installation of `gert` for `devtools` requires `libgit2-dev` on Ubuntu.

## Help files
After loading the package with `library(ntsworkflow)`, You can find documentation by typing `browseVignettes("ntsworkflow")`. 

## License

Copyright 2025 Bundesanstalt für Gewässerkunde (Federal Institute of Hydrology)

ntsworkflow is free software: you can redistribute it and/or modify it under the 
terms of the GNU General Public License as published by the Free Software 
Foundation, either version 3 of the License, or (at your option) any 
later version.

ntsworkflow is distributed in the hope that it will be useful, but WITHOUT ANY 
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS 
FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details.

You should have received a copy of the GNU General Public License along 
with ntsworkflow. If not, see <https://www.gnu.org/licenses/>.


## Change log

### 0.2.7

In `Report$integRes`, the column `samp` contains only the sample name (previously sample path)
